---
title: "Task1"
output: html_document
---
  
Загрузим наши данные:

```{r}
df <- read.table(file = 'study_fall2020_dataprog_creditcard.txt',
                sep = ';',
                header = TRUE)
head(df)
```


## Информация о dataset:

card: Factor. Was the application for a credit card accepted?


reports: Number of major derogatory reports.


age: Age in years plus twelfths of a year.


income: Yearly income (in USD 10,000).


share: Ratio of monthly credit card expenditure to yearly income.


expenditure: Average monthly credit card expenditure.


owner: Factor. Does the individual own their home?


selfemp: Factor. Is the individual self-employed?


dependents: Number of dependents.


months: Months living at current address.


majorcards: Number of major credit cards held.


active: Number of active credit accounts.


## Task:


1. Write a dataframe with the following properties:


- Let's consider 'today' is 2000-01-01. Convert amount of months applicant living at his/her current address to days


```{r}
#install.packages("lubridate")
#install.packages("dplyr")
library(lubridate)
library(dplyr)

now <- as.Date("2000-01-01")
df <- mutate(df, days = as.numeric( (now %m+% months(df$months)) - now))

head(df)
```

- Add column with income per family member (not the number of dependents!)


inPerFamMem in USD 10,000

```{r}

df <- mutate(df, inPerFamMem = df$income/(df$dependents + 1))

head(df)
```

- Define your own criterion of income level, and split data according to levels of this criterion


Для определения собственного уровня дохода воспользуемся переменной inPerFamMem - доход на члена семьи. Выдвигаем предположение о равестве средних в двух группах по фактору "selfemp". Пользуемся критерием Вилкоксана о равенстве медиа, так как в данных наблюдаются выбросы.

```{r}
wilcox.test(df$inPerFamMem[df$selfemp == 'yes'], df$inPerFamMem[df$selfemp == 'no'])

```

У нас нету оснований отвергнуть гипотезу о равенстве медиан.


Определим собственый критерий, как средний доход на члена семьи больше среднего и владение собственным домом:
```{r}
library(plyr)

df <- mutate(df, myCrit = as.factor((df$inPerFamMem > mean(df$inPerFamMem)) & (df$owner == 'yes')))
df <- mutate(df, myCrit = revalue(df$myCrit, c("FALSE"= 0, "TRUE"=1)))

head(df)

```

- Information about only self-employed people should be reported
- Use colon as a field separator


Выведем подсет данных в файл, где наш критерий = 1, selfemp = yes, разделитель - ":"
```{r}
df_ <- df[c(1:8,11:15)]
head(df_)

write.table(df_[which((df$myCrit == 1) & (df$selfemp == 'yes')),],
            file = "myData.txt",
            sep = ":")
```


2. Print some information about dataset (either to file or to standard output)


- Average amount of dependents for people don't own their home.


```{r}
mean(df$dependents[df$owner == "no"])
```


- Average age in each income groups (use grouping variable defined in 1.)
```{r}
aggregate(x = df$age,                
          by = list(df$myCrit),              
          FUN = mean)
```


- Top 5 youngest and eldest people, whose application was declined


```{r}
require(data.table)
d <- data.table(df, key="age")
d[d$card == 'no', head(.SD, 5)]
d[d$card == 'no', tail(.SD, 5)]
```


- Average number of major CCs held for people with top 10 income

```{r}
require(data.table)
d <- data.table(df, key="income")
mean(d[, tail(.SD, 10),]$majorcards)
```
