---
title: "Task1"
output: html_document
---
  
Загрузим наши данные:

```{r}
df <- read.table(file = 'study_fall2020_dataprog_creditcard.txt',
                sep = ';',
                header = TRUE)
head(df)
```


## Информация о dataset:

card: Factor. Was the application for a credit card accepted?


reports: Number of major derogatory reports.


age: Age in years plus twelfths of a year.


income: Yearly income (in USD 10,000).


share: Ratio of monthly credit card expenditure to yearly income.


expenditure: Average monthly credit card expenditure.


owner: Factor. Does the individual own their home?


selfemp: Factor. Is the individual self-employed?


dependents: Number of dependents.


months: Months living at current address.


majorcards: Number of major credit cards held.


active: Number of active credit accounts.


## Task:


1. Write a dataframe with the following properties:


- Let's consider 'today' is 2000-01-01. Convert amount of months applicant living at his/her current address to days


- Add column with income per family member (not the number of dependents!) inPerFamMem in USD 10,000
```{r}
library(lubridate)
library(dplyr)
library(plyr)

now <- as.Date("2000-01-01")
df <- mutate(df, days = as.numeric( (now %m+% months(months)) - now),
             inPerFamMem = income/(dependents + 1))

head(df)
```


- Define your own criterion of income level, and split data according to levels of this criterion


Для определения собственного уровня дохода воспользуемся переменной income. Введем критерий относительно медианы:


```{r}
library(ggplot2)

ggplot(data = df, aes(x = income)) +   #оценим распределение дохода
  geom_histogram(fill = "yellowgreen", 
                 color = "black")+ 
  geom_vline(xintercept = median(df$income), 
             color = "red",
             lty = 2)    

```

Так как наблюдается сильная несиметричность, то рассмотрим новую переменую - log(income) и относительно нее сформируем собственый критерий:


```{r}

df <- mutate(df, logIncome = log(income))

ggplot(data = df, aes(x = logIncome)) +   #оценим распределение log дохода
  geom_histogram(fill = "yellowgreen", 
                 color = "black")+ 
  geom_vline(xintercept = median(df$logIncome), 
             color = "red",
             lty = 2)  


scopes = quantile(df$logIncome, c(0.3255,0.6745))

```


Разделим по нашему критерию, где "3" - высокий достаток, "2" - средний, "1" - низкий.

```{r}

df <- df %>% mutate(myCrit = case_when(
                                  (logIncome >= scopes[[2]] ~ "3"), 
                                  (logIncome >= scopes[[1]] & logIncome < scopes[[2]] ~ "2"),
                                  (logIncome < scopes[[1]]  ~ "1"),
                                  TRUE ~ "NA"))
df <- mutate(df, myCrit = as.factor(myCrit))
head(df)
```


- Information about only self-employed people should be reported
- Use colon as a field separator


Выведем подсет данных в файл, где наш критерий >= 2, selfemp = yes, разделитель - ":"
```{r}
df_ <- df

subset <- filter(df_, as.numeric(myCrit) >= 2 &
                      selfemp == "yes"
                 )
write.table(subset, file = "myData.txt", sep = ":",
            row.names = TRUE, col.names = NA)

```


2. Print some information about dataset (either to file or to standard output)


- Average amount of dependents for people don't own their home.


```{r}
owner_no <- filter(df, owner == "no") 
summarise(owner_no, mean(dependents))
```


- Average age in each income groups (use grouping variable defined in 1.)
```{r}
by_age <- df %>% group_by(myCrit) %>% summarize_all("mean")
by_age$age
```


- Top 5 youngest and eldest people, whose application was declined


```{r}
subset <- filter(df, card == "no")
subset <- arrange(subset, age)
head(subset,5)
tail(subset,5)
```


- Average number of major CCs held for people with top 10 income

```{r}
df_ <- head(arrange(df, desc(income)), 10)
mean(df_$majorcards)

```

## Task 2

1. Опишите распределения данных по возрасту, полу и уровню дохода


К сожалению в данных нету информации про пол претендентов, поэтому опишем распределения данных только по возрасту и уровню дохода.


```{r}

ggplot(data = df, aes(x = age)) +
  geom_histogram(fill = "yellowgreen", 
                 color = "black")+ 
  geom_vline(xintercept = median(df$age), 
             color = "red",
             lty = 2)

ggplot(data = df, aes(x = income)) +
  geom_histogram(fill = "yellowgreen", 
                 color = "black")+ 
  geom_vline(xintercept = median(df$income), 
             color = "red",
             lty = 2)


ggplot(data = df, aes(y = age)) +
  geom_boxplot(fill = "yellowgreen", 
                 color = "black")

ggplot(data = df, aes(y = income)) +
  geom_boxplot(fill = "yellowgreen", 
                 color = "black")

```


2. Опишите структуру трат
Так как распределения не симметричны, прологарифмируем данные income и expenditure:

```{r}

df <- mutate(df, logIncome = log(income),
                 logExp = log(expenditure),
                 logShare = log(share))


ggplot(data = df, aes(x = logExp)) +
  geom_histogram(fill = "yellowgreen", 
                 color = "black")


ggplot(data = df, aes(x = logIncome)) +
  geom_histogram(fill = "yellowgreen", 
                 color = "black")

ggplot(data = df, aes(y = logExp)) +
  geom_boxplot(fill = "yellowgreen", 
                 color = "black")


ggplot(data = df, aes(y = logIncome)) +
  geom_boxplot(fill = "yellowgreen", 
                 color = "black")


```


```{r}
library(psych)
df_for_cor <- data.frame( "reports"    = df$reports, 
                          "age"        = df$age, 
                          "income"     = df$logIncome,
                          "share"      = df$logShare,
                          "expen"      = df$expenditure,
                          "depen"      = df$dependents,
                          "months"     = df$months,
                          "majorcards" = df$majorcards,
                          "active"     = df$active)

pairs.panels(df_for_cor, 
             method = "pearson", 
             hist.col = "orange",
             density = TRUE,
             ellipses = FALSE,
             lm = TRUE,
             stars = TRUE
             )

ggplot(data = df, aes(x = logIncome, y = logExp)) +
geom_point() + 
labs(title = "Log income VS Log expenditure", 
     x = "log income", 
     y = "log expenditure")

```


```{r}
ggplot(data = df, aes(x = logIncome, y = logShare)) +
geom_point() + 
labs(title = "Log income VS Log share", 
     x = "log income", 
     y = "log share")


ggplot(data = df, aes(x = age, y = logIncome)) +
geom_point() + 
labs(title = "Log income VS Log share", 
     x = "age", 
     y = "log income")
```


3. Опишите зависимость, выдана карта, или нет от других признаков


```{r}

df$card <- revalue(df$card, c("yes" = 2, "no" = 1))
df$card <- as.numeric(df$card)

my_cols <- c("light blue", "orange")
pairs(~ reports + logIncome + majorcards + dependents + months + logShare, data = df, col = my_cols[df$card], pch = 19,  cex = 0.5, oma=c(3,3,6,8))
```


